The \proglang{R} Package \pkg{forestinventory}: Design-Based Global and Small Area Estimations for Multiphase Forest Inventories

### 2018-11-15 2nd revision/answer_to_editor.Rnw

%------------------------------------------------------------------------------------------------%
% ---------------------------------- Answer -----------------------------------------------------%

\section*{Answer}

The reviewers code runs without bugs but the simulation was incorrectly formulated. One fundamental problem with the reviewers simulation is that he seems to pose all three examples in the \textit{finite} population context rather than the \textit{infinite} population context. We can see this right at the beginning:\\

Here, the reviewer sets up a dataframe defining a \textit{finite} population that he then uses in his simulation examples:

@@

\begin{small}
<<>>=
# Reviewer's artificial data:
set.seed(0)
d <- data.frame(
          elv = (e<-runif(500, 300, 550)),     # explanatory variable (observed in s1)
          vol1 = rnorm(e, mean=600-e, sd=100), # response variables (to be observed in s2 only)
          vol2 = rnorm(e, mean=300, sd=100)*(e<310),
          resp3 = runif(e) < 0.03)
@
\end{small}

Let's look at the first simulation example to explain where the fundamental problem is:

\begin{small}
<<eval=FALSE>>=
library(forestinventory)
# PROBLEM 1: Small sample size (n2=5)
true_value <- mean(d$vol1)
cover <- replicate(10000, {
d$phase <- 1
d$phase[sample(length(d$phase), 5)] <- 2
ci <- confint( twophase( vol1 ~ elv, data = d,
phase_id = list(phase.col="phase", terrgrid.id=2) ), level=0.95 )
ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
} )
mean(cover)
@
\end{small}

As obvious, the true population mean (\code{true\_value}) that has to be estimated is here calculated by taking the mean of a \textit{finite} number of response variable values ($N=500$, i.e. \code{nrow(d)}). This shows that the reviewer actually defined a finite population for the response variable (\code{vol1}), and likewise for the auxiliary variable (\code{elv}). In every iteration of his simulation, he then resamples from a finite set of observed response variable values $s_2$. This is a fundamental misunderstanding of the infinite population theory, which was exclusively used to derive all estimators contained in \code{forestinventory}. First of all, the true population mean (\code{true\_value}) should be an integral (i.e. spatial mean) over infinite points, as described in section 2.1 in the article, but by using the finite mean, the reviewer has effectively acknowledged that actually $N=500$ instead of infinity.\par
Second, we then see that every population unit of his finite population is assigned to the first phase (\code{d\$phase<- 1}). Basically, this means that the auxiliary information is not resampled in every simulation iteration, but the entire population ($N=500$) of auxiliary information \code{elv} is used. This indicates that the reviewer intended to apply an \textit{exhaustive} two-phase estimator where the true population mean of the auxiliary information is used. Notice that in the \code{twophase()}-function call, the reviewer, however, did \textit{not} set the exhaustive parameter as demonstrated in section 3.1 pp. 11-12 in the article, making this a non-exhaustive function call. This means that by simulation design $s_1$ does not get resampled with every replicate so the variability coming from resampling the first phase is ignored.\\

@@

All 3 examples given by the reviewer made these same errors so the results clearly can't be considered conclusive on their own merits.  This is why we attempted to reformulate the simulation examples correctly in the infinite population context and mimic the examples intended effects. The following is some sample code (similar code was added to the Appendix in the last submission) to demonstrate simulation in the infinite population context.  This technique has already applied in Mandallaz (2013a), Mandallaz et al. (2013) and Massey et al. (2015):

\begin{small}
<<eval=FALSE>>=
library(forestinventory)
library(rmutil)  # used to calculate the integral in the spatial mean
#
#
set.seed(1)
#
# --- function to create density surface: --- #
target.surface <- function(x0, y0){
  local.density <- 30 + 13*x0 - 6*y0 - 4*x0^2 + 3*x0*y0 + 2*y0^2 + 6*cos(pi*x0)*sin(pi*y0)
  local.density
}
#
# --- True spatial mean for Global Surface Area: --- #
true_value <- int2(target.surface, a = c(0,0), b=c(3,2))/6 # 39.16667  #true mean
#
# --- sample generator for simple two-phase sampling: --- #
sample.generator<- function(n1, n2){

  realization <- matrix(NA,n1,6)
  for(i in 1:n1){
    x0=2*runif(1) # NOTE: The randomness comes from the uniform independent selection of points
    y0=3*runif(1)
    realization[i,1] <- target.surface(x0, y0)
    realization[i,2] <- x0
    realization[i,3] <- y0
    realization[i,4] <- x0*x0
    realization[i,5] <- x0*y0
    realization[i,6] <- y0*y0
  }
  realization <- as.data.frame(realization)
  names(realization) <- c("response","x","y","xx","xy","yy")
  realization$phase <- 1
  realization$phase[sample(nrow(realization), n2)] <- 2
  realization$response[realization$phase == 1] <- NA
  realization
}
@
\end{small}

As obvious, we take a finite random sample $s_2$ and $s_1$ from an infinite number of points (note that the number of possible points, i.e., coordinate pairs \code{x0, y0}) as input for the density function \code{target.surface} is infinite). The randomness in the sampling procedure comes from the uniform independent selection of points that are randomly distributed over the density surface defining the distribution of the target variable at every possible point in the plane.

\subsection*{Reviewer Problem \#1}

@@

\begin{small}
<<eval=FALSE>>=
# --- Run 10000 simulations with n1 = 500 and n2 = 5 (small sample size n2) --- #
n1 <- 500
n2 <- 5
num_sims <- 10000
ci.g.logical <- rep(NA, num_sims)   # two-phase coverage indicators for g-weight formula
ci.ext.logical <- rep(NA, num_sims) # two-phase coverage indicators for external formula
ci.op.logical <- rep(NA, num_sims)  # one-phase coverage indicators
#
for(j in 1: num_sims){
  realization <- sample.generator(n1 = n1, n2 = n2)
  est <- twophase(formula = response ~ y + x + xx, data=realization, 
                  phase_id = list(phase.col = "phase", terrgrid.id = 2))
  est_onephase <- onephase(formula = response ~ 1, data=realization, 
                           phase_id = list(phase.col = "phase", terrgrid.id = 2))
  ci <- confint(est)
  ci_op <- confint(est_onephase)
  ci.g.logical[j] <- ci$ci$ci_lower_g <= true_value & true_value <= ci$ci$ci_upper_g
  ci.op.logical[j] <- ci_op$ci$ci_lower_op <= true_value & true_value <= ci_op$ci$ci_upper_op
}
#
# results for coverage rates:
mean(ci.g.logical)   # 0.6738, g-weight coverage rate of 2-phase estimator
mean(ci.op.logical)  # 0.9457, coverage rate of one-phase estimator
@
\end{small}

The one-phase estimator has the correct coverage rate even with sample size 5 (it would also have it for $n_2$ = 2) but it will probably have a very large margin of variance (i.e. estimation error). The g-weight variance has lower than nominal coverage rates because the contribution of the variance by calculating the beta vector is based on approximations (i.e. a first order taylor approximation for the g-weight) which are not adequate for this sample size.  The article has already included a graphic with sample code that addresses the sample size issue and the reviewer has already accepted it so we can leave this issue alone for the time being.

\subsection*{Reviewer Problem \#2 and \#3}

The 2nd reviewer example involves creating a response variable \code{vol2}, that is \code{0} for 98.2\% of the values and non-zero where the auxiliary information yields a low value.  Presumably this would occur in situations where there is a treeline dependent on elevation and the vast majority of the inventory domain is at elevations where no trees can grow.  This is problematic for the g-weight and external estimators because all plots in $s_2$ where trees actually exist are high leverage outliers. However, a quick look at the reported R-squared, which would be very close to zero, should immediately trigger the  question why the model is included in the first place.\\

### 2018-11-15 JStatSoft_forestinventory_master-concordance.tex
update 15_11_2018 2nd revision
update 15_11_2018 2nd revision

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
121 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
127 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
1 2 8 0 1 2 130 1 1 5 7 0 1 2 2 1 1 2 27 0 1 2 3 1 1 2 13 0 1 2 5 1 %
1 5 4 0 1 1 8 0 1 2 7 1 1 5 7 0 1 2 2 1 1 2 9 0 1 2 7 1 1 3 5 0 1 2 %
1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 5 7 0 1 2 2 1 1 2 9 0 %
1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
133}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
102 1 1 3 2 0 1 1 3 0 1 2 4 1 1 5 7 0 1 2 3 1 1 2 12 0 1 2 4 1 1 5 7 %
0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
135 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
1 2 1 1 4 38 1 1 2 11 0 1 2 80 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1309:%
141 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1314:%
1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1549:%
147 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1554:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
0 1 2 2 1 1 6 8 0 1 2 4 1 1 4 6 0 1 2 6 1 1 2 4 0 1 2 40 1 1 2 11 0 %
1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
153 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
159 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
1 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1772:%
165 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1777:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 22 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1956:%
171 11 1}

>>>>

\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:%
1 105 1 1 4 1 1 1 4 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Introduction.Rnw:ofs 113:%
1 64 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 178:%
121 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Main structure and modules of the Package.Rnw:ofs 183:%
1 67 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 251:%
127 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Twophase_Estimators and their Application.Rnw:ofs 256:%
1 2 1 1 4 107 1 1 2 1 0 2 1 3 0 1 2 2 1 1 6 13 0 1 1 1 2 16 1 1 4 6 %
0 1 2 4 1 1 2 24 0 1 2 30 1 1 3 5 0 1 2 4 1 1 2 1 0 1 5 6 0 1 2 3 1 %
1 2 8 0 1 2 130 1 1 5 7 0 1 2 2 1 1 2 27 0 1 2 3 1 1 2 13 0 1 2 5 1 %
1 5 4 0 1 1 8 0 1 2 7 1 1 5 7 0 1 2 2 1 1 2 9 0 1 2 7 1 1 3 5 0 1 2 %
1 1 1 6 5 0 1 1 3 0 1 2 1 1 1 2 11 0 1 2 3 1 1 5 7 0 1 2 2 1 1 2 9 0 %
1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 817:%
133}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Threephase_Estimators and their Application.Rnw:ofs 818:%
1 2 1 1 4 91 1 1 2 1 0 1 1 3 0 1 2 4 1 1 4 6 0 1 2 2 1 1 2 29 0 1 2 %
102 1 1 3 2 0 1 1 3 0 1 2 4 1 1 5 7 0 1 2 3 1 1 2 12 0 1 2 4 1 1 5 7 %
0 1 2 2 1 1 2 12 0 1 2 6 1 1 5 7 0 1 2 2 1 1 2 12 0 1 2 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1169:%
135 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Calculation of Confidence Intervals.Rnw:ofs 1174:%
1 2 1 1 4 38 1 1 2 11 0 1 2 80 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1309:%
141 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Special Cases and Scenarios.Rnw:ofs 1314:%
1 12 1 1 2 4 0 1 2 4 1 1 4 6 0 1 2 4 1 1 4 6 0 1 2 11 1 1 8 16 1 1 5 %
7 0 1 2 2 1 1 2 8 0 1 2 2 1 1 2 8 0 1 2 13 1 1 2 4 0 1 2 27 1 1 2 4 %
0 1 2 26 1 1 2 4 0 1 2 24 1 1 2 4 0 1 2 17 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1549:%
147 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Analysis and Visualization.Rnw:ofs 1554:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 5 7 0 1 2 2 1 1 5 7 0 1 2 2 1 1 6 8 %
0 1 2 2 1 1 6 8 0 1 2 4 1 1 4 6 0 1 2 6 1 1 2 4 0 1 2 40 1 1 2 11 0 %
1 2 16 1 1 2 4 0 1 2 11 1 1 3 5 0 1 2 5 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1743:%
153 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Future Plans.Rnw:ofs 1748:%
1 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1760:%
159 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Acknowledgements.Rnw:ofs 1765:%
1 6 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1772:%
165 4 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:./rnw/Appendix.Rnw:ofs 1777:%
1 2 1 1 4 8 1 1 4 6 0 1 2 3 1 1 3 2 0 2 1 3 0 1 2 2 1 1 24 26 0 1 2 %
3 1 1 70 71 0 1 2 6 1 1 3 2 0 2 1 1 4 6 0 1 2 11 1}
\Sconcordance{concordance:JStatSoft_forestinventory_master.tex:JStatSoft_forestinventory_master.rnw:ofs 1945:%
171 11 1}
